Differentiation By Year

### 2017-08-17 doc/RMD/by-year.Rmd
functionalize ggscatter plot

```{r dapc_plot, fig.width = 7, fig.height = 7}
RYD <- bind_cols(Population = regyear.dapc$grp, strata(dat11cc), as.data.frame(regyear.dapc$ind.coord)) %>%
  as_tibble()
RYD_pop <- RYD %>% 
  group_by(Population) %>% 
  summarize_if(is.numeric, mean, na.rm = TRUE) %>%
  rename_all(function(x) gsub("LD", "mean", x))
RYD <- full_join(RYD, RYD_pop, by = "Population")
RYD_PLOT <- ggplot(RYD, aes(x = LD1, y = LD2, color = Year)) + 
  geom_text(aes(label = Year), alpha = 0.5) +
  geom_segment(aes(x = mean1, y = mean2, xend = LD1, yend = LD2), alpha = 0.5) +
  stat_ellipse(type = "norm", level = 0.66, alpha = 0.75) + 
  facet_wrap(~Region) +
  theme_bw() +  
  theme(aspect.ratio = 0.8) +
  theme(legend.position = "bottom") +
  theme(axis.text = element_blank()) + 
  theme(axis.title = element_blank()) + 
  theme(axis.ticks = element_blank()) + 
  viridis::scale_color_viridis(discrete = TRUE, direction = -1) +
  viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
  scale_y_continuous(breaks = 0) + 
  scale_x_continuous(breaks = 0) + 
  theme(panel.background = element_rect(fill = "grey95")) +
  theme(panel.grid.major = element_line(color = "black"))
RYD_PLOT
par(mfrow = c(2, 1))
loadingplot(regyear.dapc$var.contr, axis = 1)
loadingplot(regyear.dapc$var.contr, axis = 2)
par(mfrow = c(1, 1))
```

@@

@@

ggloc <- function(dat, REG, years = 2003:2008){
  map(dat, . %>% 
    filter(Region == REG) %>%
    {ggplot(., aes(x = Year, y = frequency, color = allele, group = allele)) +
    geom_line() + 
    geom_point(aes(size = N), show.legend = FALSE) +
    # geom_label(aes(label = allele), size = 1, color = "black") +
    scale_x_continuous(breaks = years) +
    scale_y_continuous(limits = c(0, 1)) +
    theme_bw(base_size = 16, base_family = "Helvetica") +
    theme(legend.position = "top") +
    theme(aspect.ratio = 1/2) +
    theme(axis.text = element_text(color = "black")) +
    theme(axis.ticks.y = element_blank()) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    # theme(axis.text.x.top = element_text(angle = 90, vjust = 0.5, hjust = 0)) +
    theme(panel.grid.minor.x = element_line(linetype = 0, color = "grey50")) +
    theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
    theme(panel.grid.major = element_line(colour = "grey20")) +
    theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
    # theme(panel.spacing.y = unit(0, "line")) +
    theme(panel.background = element_rect(color = NA, fill = "grey96")) +
    theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
    theme(panel.border = element_blank()) +
    scale_size(range = c(3, 6)) +
    scale_color_viridis(option = "C", end = 0.9, 
                        guide = "legend",
                        breaks = sort(unique(.$allele)))
  })
}
ggl <- ggloc(loc, "WA")

@@

```{r NEplot, fig.width = 10, fig.height = 10}
gglNE <- ggloc(loc, "NE", 2003:2012)
cowplot::plot_grid(plotlist = gglNE, labels = names(gglNE), ncol = 2)
```
```{r MIplot, fig.width = 10, fig.height = 10}
gglMI <- ggloc(loc, "MI", 2003:2012)
cowplot::plot_grid(plotlist = gglMI, labels = names(gglMI), ncol = 2)
```

>>>>

```{r dapc_plot, fig.width = 7, fig.height = 7, warning = FALSE, message = FALSE}
# A ggplot2 version of scatter.dapc
#
# @param DAPC an object of class "dapc" derived from [adegenet::dapc]
# @param STRATA a data frame defining the population strata (see [adegenet::strata])
# @param color the variable in `STRATA` defining the color palette for the plot
# @param filter a "quosure" containing a filtering method to be passed to [dplyr::filter]
#
# @return
# @export
#
# @examples
# library("adegenet")
# data(microbov)
# strata(microbov) <- as.data.frame(other(microbov))
# setPop(microbov) <- ~spe/breed
# mscat <- ggscatter(dapc(microbov, n.pca = 20, n.da = 40), strata(microbov),
#                    color = "breed",
#                    filter = quo(breed %in% c("Borgou", "Zebu", "Montbeliard", "Salers")))
# mscat + facet_wrap(~coun)
# mscat + facet_wrap(~spe)
# mscat + facet_wrap(~breed)
ggscatter <- function(DAPC, STRATA, color = "Year", filter = NULL){
  RYD <- bind_cols(Population = DAPC$grp, STRATA, as.data.frame(DAPC$ind.coord)) %>%
    as_tibble()
  RYD_pop <- RYD %>% 
    group_by(Population) %>% 
    summarize_if(is.numeric, mean, na.rm = TRUE) %>%
    rename_all(function(x) gsub("LD", "mean", x))
  RYD <- full_join(RYD, RYD_pop, by = "Population")
  yminor <- pretty(RYD$LD2)
  xminor <- pretty(RYD$LD1)
  RYD <- if (!is.null(filter)) filter(RYD, !!filter) else RYD
  RYD_PLOT <- ggplot(RYD, aes_string(x = "LD1", y = "LD2", color = color)) + 
    geom_text(aes_string(label = color), alpha = 0.75) +
    geom_segment(aes(x = mean1, y = mean2, xend = LD1, yend = LD2), alpha = 0.5) +
    stat_ellipse(type = "norm", level = 0.66, alpha = 0.75) + 
    theme_bw(base_size = 16, base_family = "Helvetica") +  
    theme(aspect.ratio = 0.8) +
    theme(legend.position = "bottom") +
    theme(axis.text = element_blank()) + 
    theme(axis.title = element_blank()) + 
    theme(axis.ticks = element_blank()) + 
    viridis::scale_color_viridis(discrete = TRUE, direction = -1) +
    viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
    scale_y_continuous(breaks = 0, minor_breaks = yminor) + 
    scale_x_continuous(breaks = 0, minor_breaks = xminor) + 
    theme(panel.background = element_rect(fill = "grey95")) +
    theme(panel.grid.major = element_line(color = "black")) +
    theme(panel.grid.minor = element_line(color = "white")) 
  RYD_PLOT
}

par(mfrow = c(2, 1))
loadingplot(regyear.dapc$var.contr, axis = 1)
loadingplot(regyear.dapc$var.contr, axis = 2)
par(mfrow = c(1, 1))

ggscatter(regyear.dapc, strata(dat11cc)) +
  facet_wrap(~Region) +
  theme(legend.position = c(0.75, 0.1)) +
  guides(color = guide_legend(nrow = 4)) +
  theme(legend.direction = "horizontal")
```

@@

```{r dapc_plot_newamica, fig.width = 5, fig.height = 4, warning = FALSE, message = FALSE}
ggscatter(regyear.dapc, strata(dat11cc), filter = quo(Region %in% c("WA", "CA", "NE", "MI"))) +
  facet_wrap(~Region, nrow = 2) +
  theme(legend.position = "right") +
  theme(legend.key = element_rect(fill = "grey95"))
```

@@

ggloc <- function(dat, REG, years = 2003:2008){
  map(dat, . %>% 
    filter(Region == REG) %>% {
      ggplot(., aes(x = Year, y = frequency, color = allele, group = allele)) +
        geom_text(aes(label = paste0("(", N, ")")), y = 0.99, color = "grey20", vjust = 1) +
        geom_line(alpha = 0.75) +
        geom_point(aes(size = N), show.legend = FALSE, alpha = 0.75) +
        scale_x_continuous(breaks = years) +
        scale_y_continuous(limits = c(0, 1)) +
        theme_bw(base_size = 16, base_family = "Helvetica") +
        theme(legend.position = "top") +
        theme(aspect.ratio = 1/2) +
        theme(axis.text = element_text(color = "black")) +
        theme(axis.ticks.y = element_blank()) +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
        # theme(axis.text.x.top = element_text(angle = 90, vjust = 0.5, hjust = 0)) +
        theme(panel.grid.minor.x = element_line(linetype = 0, color = "grey50")) +
        theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
        theme(panel.grid.major = element_line(colour = "grey20")) +
        theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
        # theme(panel.spacing.y = unit(0, "line")) +
        theme(panel.background = element_rect(color = NA, fill = "grey98")) +
        theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
        theme(panel.border = element_blank()) +
        scale_size(range = c(3, 6), limits = c(6, 25)) +
        viridis::scale_color_viridis(option = "C", end = 0.9, 
                                     guide = "legend",
                                     breaks = sort(unique(.$allele)))
  })
}
ggl <- suppressMessages(ggloc(loc, "WA"))

@@

```{r NEplot, fig.width = 10, fig.height = 10}
gglNE <- suppressMessages(ggloc(loc, "NE", 2003:2012))
cowplot::plot_grid(plotlist = gglNE, labels = names(gglNE), ncol = 2)
```
```{r MIplot, fig.width = 10, fig.height = 10}
gglMI <- suppressMessages(ggloc(loc, "MI", 2003:2012))
cowplot::plot_grid(plotlist = gglMI, labels = names(gglMI), ncol = 2)
```
